---
title: "thompson_ES&T"
author: "Raissa Mendonca"
date: "9/13/2017"
output: html_document
---

Project: Thompson and Birchtree mines, Thompson, MB, Canada (Kent State University, NiPERA, Univ. of Michigan)

Samples collected for surface water, sediment and benthic invertebrates at each sampling site. Sediment and surface water samples collected in 4 sites in reference tributaries and 10 in exposure tributaries at both mine sites. Invertebrate samples collected in 10 sites in both reference and exposure tributaries at both mine sites.

##Reviewer's comments from ES&T

  1. Correlations between water chemistry and sediment geochemistry for SI
  2. Relationships between relative abundances and chemistry variables for SI
  3. RDA including Conductivity + pH + TOTNi + FeHFO in one model
  4. SEM-AVS/ƒOC EC10 without the outliers (SEM<AVS)
  5. Identify SEM-AVS/ƒOC outliers on all 3 concentration-response graphs
  6. Compare concentration-response relationships between EPH and TAL
  7. Sigmoid concentration-response model (code from 'compiled_R_tips.Rmd')

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
options(device=quartz)
#options(device="RStudioGD")
par(pin=c(2.5,2.5),las=1)
```

```{r Load datasets, include=F}
setwd("~/Documents/Kent State University/Costello Lab/GitHub/Thompson")
#setwd("~/Documents/GitHub/Thompson") #Office PC path to work directory // Git repository
library(vegan)
library(FSA)

thomp <- read.csv(file="thomp_chem.csv") # Chem variables NOT separated in .surf and .deep
  thomp$Site <- as.factor(thomp$Site)
  thomp$Depth <- factor(thomp$Depth,levels=c("Surface","Deep"),ordered=T)
  thomp$SEMAVS <- (thomp$SEMNi/58.69)-thomp$AVS
  thomp$SEMAVSfOC <- ((thomp$SEMNi/58.69)-thomp$AVS)/(thomp$C/100) # Org C corrected
  thomp$SEMAVSfOC[thomp$SEMAVSfOC<0] <- 15 #Set neg. values as minimum SEM-AVS/fOC at all Exp sites (similar in Birch/Weir)
  thomp$C[c(30,36)] <- NA # Outliers (not good numbers, take out from all analyses)
  thomp$C.avg <- c(rep((mean(thomp$C[1:14])),times=14),rep((mean(thomp$C[15:28])),times=14))
  names(thomp)

geochem <- read.csv(file="thomp_chem_surfdeep.csv") # Chem variables .surf and .deep, only 4 ref values

allrefs <- read.csv(file="thomp_chem_allrefs.csv") # Chem variables .surf and .deep, ref values are averages
  allrefs$SEMAVSfOC.surf[ allrefs$SEMAVSfOC.surf<0] <- 15 # Make all values min positive value from exposure sites
  allrefs$SEMAVS.surf[ allrefs$SEMAVS.surf<0] <- 0.320 # Make all values min positive value from exposure sites
  allrefs$Ni.Fe <- allrefs$TotalNi.surf-allrefs$ASCNi.surf # Variable of Ni corrected for Fe, i.e., portion of total nickel not bound to oxides

benthos <- read.csv(file="thomp_benthos.csv") # Only benthic inverts (excluded cladocera, water mites and fish)
  benthos$OL <- NULL
benthos.exp <- benthos[c(11:20,31:40),]
  
arealdens <- read.csv(file="thomp_arealdensity.csv") # Invert areal density (#/m2)
arealdens$OL <- NULL

biomass <- read.csv(file="thomp_hexbiomass.csv")
biomass$Site <- as.factor(biomass$Site)
biomass[24,] = NA
summary(biomass)

thomp.exp <- subset(allrefs,Effluent=="Exposure") # Chem variables .surf and .deep, thomp.exp only
  thomp.exp$C.avg <- c(rep((mean(thomp.exp$C.surf[1:10])),times=10),rep((mean(thomp.exp$C.surf[11:20])),times=10))
  thomp.exp$NiSEMHFO <- thomp.exp$SEMNi.surf-thomp.exp$ASCNi.surf
  thomp.exp$NiSEMDIT <- thomp.exp$SEMNi.surf-thomp.exp$DITNi.surf
  thomp.exp$NiSEMHFOfOC <- ((thomp.exp$SEMNi.surf/58.69)-(thomp.exp$ASCNi.surf/58.69))/(thomp.exp$C.avg/100) #Molar scale
  thomp.exp$SEMAVSfOC.avg = ((thomp.exp$SEMNi.surf/58.69)-thomp.exp$AVS.surf)/(thomp.exp$C.avg/100) # Org C corrected
  thomp.exp$SEMAVSfOC.avg[thomp.exp$SEMAVSfOC.avg<0] = 7.130744 # Make all values > 0 to make graph on log scale; 7.130744 is min positive value from exposure sites

chem.inv <- read.csv(file="thomp_chem_inverts.csv") # Chem variables .surf and .deep, ref values are averages, TOT_AB = total abund, REL_EPH = Ephemeroptera rel abund
  chem.inv$OL <- NULL
  chem.inv$C.avg <- c(rep(thomp$C.avg[1],20),rep(thomp$C.avg[15],20))  
  chem.inv <- cbind(chem.inv[,1:65],allrefs$Ni.Fe,chem.inv[,66:106])
  names(chem.inv)[66] <- "Ni.Fe"
  chem.inv$REL_EPH <- chem.inv$REL_EPH*100
  chem.inv$SEMAVSfOC.avg <- ((chem.inv$SEMNi.surf/58.69)-chem.inv$AVS.surf)/(chem.inv$C.avg/100)
  chem.inv$SEMAVSfOC.avg[chem.inv$SEMAVSfOC.avg<0] = 4.898928 # Make all values > 0 to make graph on log scale; 4.898928 is min positive value from ref and exp sites
  chem.inv$NiSEMHFOfOC <- ((chem.inv$SEMNi.surf/58.69)-(chem.inv$ASCNi.surf/58.69))/(chem.inv$C.avg/100) #Molar scale
  chem.inv$REL_TAL <- round((benthos$TAL/rowSums(benthos[2:39]))*100)
  chem.inv$SEMAVSfOC.surf[chem.inv$SEMAVSfOC.surf<0] <- 15
  chem.inv$SEMAVS.surf[chem.inv$SEMAVS.surf<0] <- 0.320
  
chem.inv.exp <- subset(chem.inv,Effluent=="Exposure")
  chem.inv.exp$NiSEMHFO <- thomp.exp$NiSEMHFO
  chem.inv.exp$NiSEMDIT <- thomp.exp$NiSEMDIT
  chem.inv.exp$NiSEMHFOfOC <- thomp.exp$NiSEMHFOfOC
  chem.inv.exp$SEMAVSfOC.avg <- thomp.exp$SEMAVSfOC.avg  
  chem.inv.exp$REL_TAL <- round((benthos.exp$TAL/rowSums(benthos.exp[2:39]))*100)
  REL_EPH.avg <- with(chem.inv,tapply(REL_EPH,list(chem.inv$Effluent),mean))
  #REL_EPH.exp <- mean(chem.inv.exp[c(3,4,7,14,19,20),"REL_EPH"])  #Mean EPH rel abund. in exp sites closest to the mouth of the tributary ("less contaminated")
    REL_EPH.exp <- 43 #Eph rel abund at BE4 and WE10, only two exp sites within mean +/- s.d. or s.e. of background [Ni] (mean at ref sites); both sites had same Eph rel abund of 0.43
    REL_EPH.BE <- mean(chem.inv.exp[c(3,4,7),"REL_EPH"])
    REL_EPH.WE <- mean(chem.inv.exp[c(14,19,20),"REL_EPH"])

```

```{r Subsets and kd, include=F}
Birch <- subset(thomp,thomp$Location=="Birchtree")
Weir <- subset(thomp,thomp$Location=="Weir")
BR <- subset(Birch,Birch$Effluent=="Reference")
WR <- subset(Weir,Weir$Effluent=="Reference")
BE <- subset(Birch,Birch$Effluent=="Exposure")
WE <- subset(Weir,Weir$Effluent=="Exposure")

surf <- subset(thomp[,c(1:31,41:42)],thomp$Depth=="Surface") # Surface sediment variables
deep <- subset(thomp[,c(1:31,41:42)],thomp$Depth=="Deep") # Deep sediment variables
water <- thomp[c(1:28),c(1:4,32:40)] # water related variables

birch.kd <- mean(log10(((surf$TotalNi[surf$Location=="Birchtree"])*1000)/water$Ni[water$Location=="Birchtree"]))
weir.kd <- mean(log10(((surf$TotalNi[surf$Location=="Weir"])*1000)/water$Ni[water$Location=="Weir"]))

BE.kd <- mean(log10(((BE$TotalNi[BE$Depth=="Surface"])*1000)/BE$Ni[BE$Depth=="Surface"]))
WE.kd <- mean(log10(((WE$TotalNi[WE$Depth=="Surface"])*1000)/WE$Ni[WE$Depth=="Surface"]))
```

```{r Log axis function, include=F}
# Graphing function: log axis
logaxis <- function(minlog,maxlog,side){
  pow <- seq(minlog,maxlog,by=1)
  ticksat <- as.vector(sapply(pow, function(p) (1:10)*10^p))
  axis(side, 10^pow,las=1)
  axis(side, ticksat, labels=NA, tcl=-0.25, lwd=0, lwd.ticks=1)
}
```

```{r ECXXcalc, include=F}
ECXX=function(model,percent){
  ECname = paste("EC",percent)
  dose=names(model$dataClasses)
  EC = (100-percent)/100 
  ref = model$m$getPar()[1]
  min.x = 1
  max.x = 10000 #Might need to change the range
  range = seq(min.x,max.x,by=1)
  new.range = data.frame(frame=seq(min.x,max.x,by=1))
  names(new.range)[[1]]=dose
  bf = predict(model,newdata=new.range)
  locEC=which.min(abs(bf-ref*EC))
  se = sqrt(apply(attr(predict(model, newdata=new.range),"gradient"),1,function(x) sum(vcov(model)*outer(x,x))))
  LCI=bf+se*qnorm(0.025)
  locL=which.min(abs(LCI-ref*EC))
  UCI=bf+se*qnorm(0.975)
  locU=which.min(abs(UCI-ref*EC))
  result=c(range[locEC],range[locL],range[locU])
  result = signif(result,digits=3)
  names(result)[[1]]=ECname
  names(result)[[2]]="Lower-bound (2.5%)"
  names(result)[[3]]="Upper-bound (97.5%)"
  print(result)
  mat=cbind(new.range,bf,LCI,UCI)
  invisible(list(ResultTable=result,CImat=mat))
  	}
```

```{r ANOVA eq, include=F}
  #Get ANOVA results for full equation model
ANOVA_F=function(mod,y){
	ypred=predict(mod)
  y.no.na = y[!is.na(y)]
  dfnum = length(mod$m$getPar())-1
  dfden = length(y.no.na)-length(mod$m$getPar())
  SSerr=sum((ypred-y.no.na)^2)
	SSreg=sum((ypred-mean(y.no.na))^2)
	SStot = SSerr + SSreg
  Frat = (SSreg/dfnum)/(SSerr/dfden)
  probF = pf(Frat,dfnum,dfden,lower.tail=F)
  Rsquare = 1-(SSerr/SStot)
  result=c(Frat,probF,Rsquare)
  names(result)[[1]]="F ratio"
	names(result)[[2]]="P-value"
  names(result)[[3]]="R-squared"
  return(result)
}

```
_____

####1. WATER-SED correlations || Exposure sites only (reviewer's question about "along Ni gradient")

  a. Table of correlation coefficients between surface (rows 1-15) and deep (rows 16-30) sediment variables and water chemistry variables || Bold green cells represent significant correlations (p < 0.01)
  b. Table of correlation coefficients between surface sediment variables || Bold green cells represent significant correlations
  c. Table of correlation coefficients between water chemistry variables || Bold green cells represent significant correlations
  
```{r 1. WATER-SED correlations, echo=F}
library('formattable')
  p_bold <- formatter("span", style = x ~ style("font-weight" = ifelse(x < 0.01, "bold", NA)))
  
  ##Exposure sites only
surf = surf[c(5:14,19:28),c(6:18,21,22)] 
deep = deep[c(5:14,19:28),c(6:18,21,22)]
water = water [c(5:14,19:28),5:12]

##Set up matrices
watersurf.cor <- matrix(data=NA,ncol=8,nrow=15,dimnames=list(colnames(surf),colnames(water)))
watersurf.p <- matrix(data=NA,ncol=8,nrow=15,dimnames=list(colnames(surf),colnames(water)))
waterdeep.cor <- matrix(data=NA,ncol=8,nrow=15,dimnames=list(colnames(deep),colnames(water)))
waterdeep.p <- matrix(data=NA,ncol=8,nrow=15,dimnames=list(colnames(surf),colnames(water)))

##Water-surf sed. correlations loop
for(i in 1:ncol(surf)) {
  for(j in 1:ncol(water)) {
    cor <- cor.test(surf[,i],water[,j])
    watersurf.cor[i,j] = signif(cor$estimate,3)
    watersurf.p[i,j] = signif(cor$p.value,3)
  }
}

##Water-deep sed. correlations loop
for(i in 1:ncol(deep)) {
  for(j in 1:ncol(water)) {
    cor <- cor.test(deep[,i],water[,j])
    waterdeep.cor[i,j] = signif(cor$estimate,3)
    waterdeep.p[i,j] = signif(cor$p.value,3)
  }
}

watersed.cor = rbind(watersurf.cor,waterdeep.cor)
watersed.p = rbind(watersurf.p,waterdeep.p)


  ##P-values for water-sed correlations
  #formattable(as.data.frame(watersed.p, check.names=F), list(~ p_bold))

  ##Signif coef for water-sed correlations
  highcor <- formatter("span",style = x ~ style(color = ifelse(watersed.p < 0.01, "green", "black"),
                                                "font-weight" = ifelse(watersed.p < 0.01, "bold", NA)))  
  formattable(as.data.frame(watersed.cor, check.names=F), list(~ highcor))

##Set up matrices
surf.cor <- matrix(data=NA,ncol=15,nrow=15,dimnames=list(colnames(surf),colnames(surf)))
surf.p <-   matrix(data=NA,ncol=15,nrow=15,dimnames=list(colnames(surf),colnames(surf)))
water.cor <- matrix(data=NA,ncol=8,nrow=8,dimnames=list(colnames(water),colnames(water)))
water.p <-   matrix(data=NA,ncol=8,nrow=8,dimnames=list(colnames(water),colnames(water)))

##Surf-surf correlations loop
for(i in 1:ncol(surf)) {
  for(j in 1:ncol(surf)) {
    cor <- cor.test(surf[,i],surf[,j])
    surf.cor[i,j] = signif(cor$estimate,3)
    surf.p[i,j] = signif(cor$p.value,1)
  }
}

##Water-water correlations loop
for(i in 1:ncol(water)) {
  for(j in 1:ncol(water)) {
    cor <- cor.test(water[,i],water[,j])
    water.cor[i,j] = signif(cor$estimate,3)
    water.p[i,j] = signif(cor$p.value,3)
  }
}

  #P-values for surf-surf correlations
  #formattable(as.data.frame(surf.p, check.names=F), list(~ p_bold))

  #Coef for surf-surf correlations
  highcor <- formatter("span",style = x ~ style(color = ifelse(surf.p < 0.01, "green", "black"),
                                                "font-weight" = ifelse(surf.p < 0.01, "bold", NA)))
  formattable(as.data.frame(surf.cor, check.names=F), list(~ highcor))

  #P-values for water-water correlations
  #formattable(as.data.frame(water.p, check.names=F), list(~ p_bold))

  #Coef for water-water correlations
  highcor <- formatter("span",style = x ~ style(color = ifelse(water.p < 0.01, "green", "black"),
                                                "font-weight" = ifelse(water.p < 0.01, "bold", NA)))
  formattable(as.data.frame(water.cor, check.names=F), list(~ highcor))
```  

_____

##### *Linear relationship between NiSEM and conductivity*
  - Two Birchtree mine sites with high conductivity (> 4000 µS/cm): when included, regression not significant (p = 0.7) but when removed (as "outliers"), regression is significant (p= 0.04) with a positive relationship between NiSEM and conductivity
```{r SEMNi~Cond relationship, echo=F, fig.align='center',tidy=TRUE, results="hide"}  
##Relationship between Total Ni and Cond w/ and w/o Cond sites > 4000 (n=2)
  par(pin=c(3,3))
  plot(SEMNi.surf~Cond,chem.inv.exp)
  #abline(lm(SEMNi.surf~Cond,chem.inv.exp[chem.inv.exp$Cond<4000,]),col=2) #Best-fit line for model without Cond>4000
  summary(lm(SEMNi.surf~Cond,chem.inv.exp))
  
  plot(SEMNi.surf~Cond,chem.inv.exp[chem.inv.exp$Cond<4000,])
  abline(lm(SEMNi.surf~Cond,chem.inv.exp[chem.inv.exp$Cond<4000,]),col=2)
  summary(lm(SEMNi.surf~Cond,chem.inv.exp[chem.inv.exp$Cond<4000,]))
```

_____

####2. REL_ABUND-CHEM correlations

  - Table of correlation coefficients between relative abundances and all physicochemical and geochemical variables
  
```{r 2. REL_ABUND-CHEM correlations, echo=F}
r_ab <- read.csv("thomp_relabund_C.csv") #Relative abundances of all families by site
r_ab <- r_ab[,2:39]
allrefs <- allrefs[,4:65]

r_ab.cor <- matrix(data=NA,ncol=62,nrow=38,dimnames=list(colnames(r_ab),colnames(allrefs)))
r_ab.p <- matrix(data=NA,ncol=62,nrow=38,dimnames=list(colnames(r_ab),colnames(allrefs)))


##Rel. abund-chemistry correlations loop

  #Cor values
for(i in 1:ncol(r_ab)) {
  for(j in 1:ncol(allrefs)) {
    cor <- cor.test(r_ab[,i],allrefs[,j])
    r_ab.cor[i,j] = signif(cor$estimate,2)
  }
}

  #p-values
for(i in 1:ncol(r_ab)) {
  for(j in 1:ncol(allrefs)) {
    cor <- cor.test(r_ab[,i],allrefs[,j])
    r_ab.p[i,j] = signif(cor$p.value,3)
  }
}


library('formattable')
  #P-values for relabund-chem correlations
  p_bold <- formatter("span", 
  style = x ~ style("font-weight" = ifelse(x < 0.01, "bold", NA)))
  #formattable(as.data.frame(r_ab.p, check.names=F), list(~ p_bold))

  #Coef for relabund-chem correlations
  highcor <- formatter("span", 
  style = x ~ style(color = ifelse(r_ab.p < 0.01, "green", "black"),"font-weight" = ifelse(r_ab.p < 0.01, "bold", NA)))
  formattable(as.data.frame(r_ab.cor, check.names=F), list(~ highcor))  
```
_____

####3. WATER+SED RDA model

  - RDA model with forward-selected water and surface sediment variables. R squared increased ~5% and overall pattern is similar to separate models. 

  - Different iterations of the full model (not shown: 1. All water variables + all surface sediment variables; 2. Conductivity + pH + all surface sediment variables) either did not significantly explain community variation (1. p = 0.15) or produced a parsimonious (post-forward stepping) model identical to RDA*SURF* (2).

```{r 3. WATER+SED RDA model, echo=F, fig.align='center', tidy=TRUE}
library(vegan)

chem.inv.hel <- cbind(chem.inv[,1:65],decostand(chem.inv[,67:104],"hellinger"))
inv.hel <- chem.inv.hel[,66:103]

rda.water.base=rda(inv.hel~1,allrefs)

rda.water.sed=rda(inv.hel~TotalNi.surf+PAmFe.surf+Cond+pH,chem.inv.hel)
  #anova(rda.water.sed)
  #summary(rda.water.sed)
  RsquareAdj(rda.water.sed)
  #good.watersed = signif((goodness(rda.water.sed))*100,3)
par(pin=c(3,3))
  plot(rda.water.sed, choices = c(1,2))
  plot(rda.water.sed, choices = c(3,4)) 
   
#rda.water.sed.for=ordiR2step(rda.water.base, rda.water.sed, direction="forward") 
  #anova(rda.water.sed.for)
  #summary(rda.water.sed.for)
  #RsquareAdj(rda.water.sed.for)
  #good.watersed.for = signif((goodness(rda.water.sed.for))*100,3)
  #plot(rda.water.sed.for)  

library('formattable')  
  highgood <- formatter("span", 
  style = x ~ style(color = ifelse((x > 20), "green", "black"),"font-weight" = ifelse((x > 20), "bold", NA)))
#formattable(as.data.frame(good.watersed, check.names=F), list(~ highgood))
```

_____

##### *Log-linear relationship between EPH relative abundance and conductivity*
  - Log-linear relationship between Ephemeridae relative abundance and conductivity significant (p = 0.003), but no effect of mine location (same slope for Birchtree and Thompson mines)
  - EPH~Cond graph to be added to SI as requested by reviewer 1
```{r Log-linear rel EPH~Cond, echo=F, fig.align='center', tidy=TRUE, results="hide"}
eph.cond <- lm(REL_EPH~log(Cond),chem.inv.exp)
  summary(eph.cond)  
  #par(mfrow=c(1,2))
  #plot(eph.cond, which=c(1,2))
  #par(mfrow=c(1,1))
  #shapiro.test(resid(eph.cond))
  
  #Rough graph
    #plot(REL_EPH~Cond,chem.inv.exp[1:10,], log="x",col=2, lwd=2)
    #points(REL_EPH~Cond,chem.inv.exp[11:20,], col=1, lwd=2)
    #abline(lm(REL_EPH~log10(Cond),chem.inv.exp[1:10,]),col=2, lwd=2)
    #abline(lm(REL_EPH~log10(Cond),chem.inv.exp[11:20,]),col=1, lwd=2)
    #abline(lm(REL_EPH~log10(Cond),chem.inv.exp),col=3, lwd=2, lty=3)
    #legend('topright',c("Birch","Thompson"), text.col=c(2,1), bty="n", text.font=2)
  
  #Pretty graph
    par(pin=c(3,3))
    plot(REL_EPH~Cond, data=chem.inv.exp,log="x",pch=c(rep(16,10),rep(1,10)),lwd=3,las=1,cex=1.5,
      ylim=c(0,50), xlim=c(100,10000), col=c(rep("tan1",10),rep("dodgerblue4",10)),
      xlab=expression(paste("Conductivity (µS cm"^"-1",")")),ylab="Ephemeridae abundance (%)",xaxt="n")
    logaxis(2,4,1)

    #abline(lm(REL_EPH[1:10]~log10(Cond[1:10]), data=chem.inv.exp),lty=3,lwd=3,col="tan1")
    #abline(lm(REL_EPH[11:20]~log10(Cond[11:20]), data=chem.inv.exp),lty=1,lwd=3,col="dodgerblue4")
    abline(lm(REL_EPH~log10(Cond), data=chem.inv.exp),lty=3,lwd=3,col="darkgray")
    par(xpd=T)
    legend(13000,52,"Birchtree",col = "tan1",pt.lwd=3,bty="n",pch=16,text.font=2,pt.cex=1.5,cex=0.8)
    legend(13000,48.5,"Thompson",col = "dodgerblue4",pt.lwd=3,bty="n",pch=1,text.font=2,pt.cex=1.5,cex=0.8)
    par(xpd=F)

tal.cond <- lm(REL_TAL~log(Cond),chem.inv.exp)
  #summary(tal.cond)  
  #par(mfrow=c(1,2))
  #plot(tal.cond, which=c(1,2))
  #par(mfrow=c(1,1))
  #shapiro.test(resid(tal.cond))
  
  #plot(REL_TAL~Cond,chem.inv.exp[1:10,], log="x",col=2, lwd=2)
  #points(REL_TAL~Cond,chem.inv.exp[11:20,], col=1, lwd=2)
  #abline(lm(REL_TAL~log10(Cond),chem.inv.exp[1:10,]),col=2, lwd=2)
  #abline(lm(REL_TAL~log10(Cond),chem.inv.exp[11:20,]),col=1, lwd=2)
  #abline(lm(REL_TAL~log10(Cond),chem.inv.exp),col=3, lwd=2, lty=3)
  #legend('topleft',c("Birch","Thompson"), text.col=c(2,1), bty="n", text.font=2)  

  
```

_____

####4. SEM-AVS model w/o 'outliers'

  - Removed BE2 and WE2 data points which were <0 and calculated EC10 and EC20 for the data without the outliers. ECXX were well below the 130 umol/g benchmark (USEPA 2005).

```{r 4. SEM-AVS model w/o outliers, echo=F, fig.align='center',tidy=TRUE, results="hide"}
par(pin=c(1.5,3))

##SEM-AVS/ƒOC (avg)
plot(REL_EPH~SEMAVSfOC.avg, data=chem.inv.exp[c(1,3:11,13:20),],log="x",pch=c(rep(16,9),rep(1,9)),lwd=3,las=1,cex=1.5,
     ylim=c(0,50),xlim=c(5,1000),col=c(rep("tan1",9),rep("dodgerblue4",9)),
     xlab=expression(paste("(Ni"[SEM]*"-AVS)/",ƒ,""[OC]*" (µmol g"^"-1",")")),
     ylab="",yaxt="n", xaxt="n")
logaxis(1,3,1)
axis(2,at=c(0,10,20,30,40,50),labels=c("","","","","",""),tick=T)

abline(lm(REL_EPH[c(1,3:10)]~log10(SEMAVSfOC.avg[c(1,3:10)]), data=chem.inv.exp),lty=3,lwd=3,col="tan1")
abline(lm(REL_EPH[c(11,13:20)]~log10(SEMAVSfOC.avg[c(11,13:20)]), data=chem.inv.exp),lty=1,lwd=3,col="dodgerblue4")

#plot(REL_EPH~log(SEMAVSfOC.avg), data=chem.inv.exp[c(1,3:11,13:20),],col=as.numeric(as.factor(Location)))
AVSOCmod <- lm(REL_EPH~log(SEMAVSfOC.avg)*Location, data=chem.inv.exp[c(1,3:11,13:20),])
summary.aov(AVSOCmod)

AVSOCW <- lm(REL_EPH~log(SEMAVSfOC.avg),data=chem.inv.exp[c(11,13:20),])
summary(AVSOCW)
#abline(AVSOCW,col=2)
exp(((REL_EPH.exp*.9)-AVSOCW$coef[1])/AVSOCW$coef[2])
  #Weir EC10 <- 69 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.avg[2]
  #Weir EC10 <- 89 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp
  #Weir EC10 <- 11 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp(43)
  #Weir EC20 <- 15 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp(43)

AVSOCB <- lm(REL_EPH~log(SEMAVSfOC.avg),data=chem.inv.exp[c(1,3:10),])
summary(AVSOCB)
#abline(AVSOCB)
exp(((REL_EPH.exp*.9)-AVSOCB$coef[1])/AVSOCB$coef[2])
  #Birchtree EC10 <- 33 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.avg[2]
  #Birchtree EC10 <- 39 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp
  #Birchtree EC10 <- 9 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp(43)
  #Birchtree EC20 <- 12 umol/g dw SEM-AVS/fOC(avg) for REL_EPH.exp(43)

abs(69-33)/mean(c(69,33)) #RD.avg <- 71%
abs(89-39)/mean(c(89,39)) #RD.exp <- 78%
abs(11-9)/mean(c(11,9)) #RD.exp(43) <- 20%
abs(15-12)/mean(c(15,12)) #RD.EC20.exp(43) <- 22%

```

|Model|EC20|RD.EC20|EC10|RD.EC10|
|:-:|:-:|:-:|:-:|:-:|
|Birchtree|12 umol/g|0.22|9 umol/g|0.20|
|Thompson|15 umol/g| |11 umol/g| |
_____

####5. SEM-AVS outliers on all CR graphs

```{r 5. SEM-AVS outliers on all CR graphs, echo=F, fig.align='center',tidy=TRUE, results="hide"}
par(pin=c(1.5,3))
chem.inv.exp$SEMNi.surf.mol = chem.inv.exp$SEMNi.surf/58.69

  ##SEM Ni
plot(REL_EPH~SEMNi.surf.mol, data=chem.inv.exp[c(1,3:11,13:20),],log="x",pch=c(rep(16,9),rep(1,9)),lwd=3,las=1,cex=1.5,
     ylim=c(0,50), xlim=c(0.1,10), col=c(rep("tan1",9),rep("dodgerblue4",9)),
     xlab=expression(paste("Ni"[SEM]*" (µmol g"^"-1"," "[DW]*")")),ylab="Ephemeridae abundance (%)",xaxt="n")
logaxis(-1,1,1)
points(3.869825,14,pch=3,lwd=3,col="tan1") #Birchtree data point that was negative before conversion to min value (BE2)
points(3.879366,3,pch=3,lwd=3,col="dodgerblue4") #Weir data point that was negative before conversion to min value (WE2)

abline(lm(REL_EPH[1:10]~log10(SEMNi.surf.mol[1:10]), data=chem.inv.exp),lty=3,lwd=3,col="tan1")
abline(lm(REL_EPH[11:20]~log10(SEMNi.surf.mol[11:20]), data=chem.inv.exp),lty=1,lwd=3,col="dodgerblue4")


  ##SEM-AVS/ƒOC (avg)
plot(REL_EPH~SEMAVSfOC.avg, data=chem.inv.exp[c(1,3:11,13:20),],log="x",pch=c(rep(16,9),rep(1,9)),lwd=3,las=1,cex=1.5,
     ylim=c(0,50),xlim=c(5,1000),col=c(rep("tan1",9),rep("dodgerblue4",9)),
     xlab=expression(paste("(Ni"[SEM]*"-AVS)/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),
     ylab="",yaxt="n", xaxt="n")
logaxis(1,3,1)
axis(2,at=c(0,10,20,30,40,50),labels=c("","","","","",""),tick=T)
points(7.130744,14,pch=3,lwd=3,col="tan1") #Birchtree data point that was negative before conversion to min value (BE2)
points(7.130744,3,pch=3,lwd=3,col="dodgerblue4") #Weir data point that was negative before conversion to min value (WE2)

abline(lm(REL_EPH[1:10]~log10(SEMAVSfOC.avg[1:10]), data=chem.inv.exp),lty=3,lwd=3,col="tan1")
abline(lm(REL_EPH[11:20]~log10(SEMAVSfOC.avg[11:20]), data=chem.inv.exp),lty=1,lwd=3,col="dodgerblue4")


  ##SEM-HFO/ƒOC (avg)
plot(REL_EPH~NiSEMHFOfOC, data=chem.inv.exp[c(1,3:11,13:20),],pch=c(rep(16,9),rep(1,9)),log="x",lwd=3,las=1,cex=1.5,
     col=c(rep("tan1",9),rep("dodgerblue4",9)), xlim=c(1,100), ylim=c(0,50),
     xlab=expression(paste("(Ni"[SEM]*"-Ni"[HFO]*")/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),
     ylab="",yaxt="n",xaxt="n")
axis(2,at=c(0,10,20,30,40,50),labels=c("","","","","",""),tick=T)
logaxis(0,2,1)
points(62.17061,14,pch=3,lwd=3,col="tan1") #Birchtree data point that was negative before conversion to min value (BE2)
points(43.66952,3,pch=3,lwd=3,col="dodgerblue4") #Weir data point that was negative before conversion to min value (WE2)

abline(lm(REL_EPH[1:10]~log10(NiSEMHFOfOC[1:10]), data=chem.inv.exp),lty=3,lwd=3,col="tan1")
abline(lm(REL_EPH[11:20]~log10(NiSEMHFOfOC[11:20]), data=chem.inv.exp),lty=1,lwd=3,col="dodgerblue4")

par(xpd=T)
legend(150,52,"Birchtree",col = "tan1",lty=3,lwd=3,bty="n",pch=16,text.font=2,pt.cex=1.5,cex=0.8)
legend(150,48.5,"Thompson",col = "dodgerblue4",lty=1,lwd=3,bty="n",pch=1,text.font=2,pt.cex=1.5,cex=0.8)
par(xpd=F)

  
  ##EC20s for all three regression models

##SEMNi (molar)
  SNW <- lm(REL_EPH~log(SEMNi.surf.mol),data=chem.inv.exp[11:20,])
  summary(SNW)
  exp(((REL_EPH.exp*.8)-SNW$coef[1])/SNW$coef[2])
    #Weir EC20 <- 3.7 µmol/g SEMNi.surf.mol for REL_EPH.avg[2]
    #Weir EC20 <- 4.5 µmol/g SEMNi.surf.mol for REL_EPH.exp
    #Weir EC20 <- 0.8 µmol/g SEMNi.surf.mol for REL_EPH.exp(43)

  SNB <- lm(REL_EPH[1:10]~log(SEMNi.surf.mol),data=chem.inv.exp[1:10,])
  summary(SNB)
  exp(((REL_EPH.exp*.8)-SNB$coef[1])/SNB$coef[2])
    #Birchtree EC20 <- 1.9 µmol/g SEMNi.surf.mol for REL_EPH.avg[2]
    #Birchtree EC20 <- 2.2 µmol/g SEMNi.surf.mol for REL_EPH.exp
    #Bircthree EC20 <- 0.6 µmol/g SEMNi.surf.mol for REL_EPH.exp(43)

  abs(3.7-1.9)/mean(c(3.7,1.9)) #RD.EC20.avg <- 64%
  abs(4.5-2.2)/mean(c(4.5,2.2)) #RD.EC20.exp <- 69%
  abs(0.8-0.6)/mean(c(0.8,0.6)) #RD.EC20.exp(43) <- 28%


##SEM-AVS (molar)
  AVSOCW <- lm(REL_EPH~log(SEMAVSfOC.avg),data=chem.inv.exp[11:20,])
  summary(AVSOCW)
  exp(((REL_EPH.exp*.8)-AVSOCW$coef[1])/AVSOCW$coef[2])
    #Weir EC20 <- 79 µmol/g SEM-AVS/fOC(avg) for REL_EPH.avg[2]
    #Weir EC20 <- 162 µmol/g SEM-AVS/fOC(avg) for REL_EPH.exp  
    #Weir EC20 <- 0.4 umol/g SEM-AVS/fOC(avg) for REL_EPH.exp(43)

  AVSOCB <- lm(REL_EPH[1:10]~log(SEMAVSfOC.avg),data=chem.inv.exp[1:10,])
  summary(AVSOCB)
  exp(((REL_EPH.exp*.8)-AVSOCB$coef[1])/AVSOCB$coef[2])
    #Birchtree EC20 <- 29 µmol/g SEM-AVS/fOC(avg) for REL_EPH.avg[2]
    #Birchtree EC20 <- 37 µmol/g SEM-AVS/fOC(avg) for REL_EPH.exp  
    #Birchtree EC20 <- 4.9 umol/g SEM-AVS/fOC(avg) for REL_EPH.exp(43)

  abs(79-29)/mean(c(79,29)) #RD.EC20.avg <- 93%
  abs(162-37)/mean(c(162,37)) #RD.EC20.exp <- 126%
  abs(0.4-4.9)/mean(c(0.4,4.9)) #RD.EC20.exp(43) <- 170%


##SEM-HFO (molar)
  TNW <- lm(REL_EPH~log(NiSEMHFOfOC),data=chem.inv.exp[11:20,])
  summary(TNW)
  exp(((REL_EPH.exp*.8)-TNW$coef[1])/TNW$coef[2])
    #Weir EC20 <- 38.2 µmol/g NiSEM-HFO/fOC for REL_EPH.avg[2]
    #Weir EC20 <- 48.9 µmol/g NiSEM-HFO/fOC for REL_EPH.exp
    #Weir EC20 <- 6.3 µmol/g NiSEM-HFO/fOC for REL_EPH.exp(43)

  TNB <- lm(REL_EPH[1:10]~log(NiSEMHFOfOC),data=chem.inv.exp[1:10,])
  summary(TNB)
  exp(((REL_EPH.avg[2]*.8)-TNB$coef[1])/TNB$coef[2])
    #Birchtree EC20 <- 23.7 µmol/g NiSEM-HFO/fOC for REL_EPH.avg[2]
    #Birchtree EC20 <- 29.5 µmol/g NiSEM-HFO/fOC for REL_EPH.exp
    #Birchtree EC20 <- 4.8 µmol/g NiSEM-HFO/fOC for REL_EPH.exp(43)

  abs(38.2-23.7)/mean(c(38.2,23.7)) #RD.EC20.avg <- 47%
  abs(48.9-29.5)/mean(c(48.9,29.5)) #RD.EC20.exp <- 49%
  abs(6.3-4.8)/mean(c(6.3,4.8)) #RD.EC20.exp(43) <- 27%
```

|Model|Birch EC20 (µmol/g)|Thompson EC20 (µmol/g)|RD.EC20|
|:-:|:-:|:-:|:-:|
|SEM Ni|0.6|0.8|0.28|
|SEM-AVS/ƒOC (w/ 'outliers')|4.9|0.4|1.70|
|SEM-AVS/ƒOC (w/o 'outliers')|12|15|0.22|
|SEM-HFO/ƒOC|4.8|6.3|0.27|

_____

####6. CRM TAL

  - Cannot fit SSlogis function to any of the models
  - Linear models all p ~ 1
  - Reliable CRMs could not be developed for Talitridae
  
```{r 6. CRM TAL, echo=F, fig.align='center',tidy=T}

chem.inv$TotalNi.surf.mol = chem.inv$TotalNi.surf/58.69
chem.inv$SEMNi.surf.mol = chem.inv$SEMNi.surf/58.69

tal.ni <- lm(REL_TAL~log(TotalNi.surf),data=chem.inv) #Can also separate analysis by [1:20,] and [21:40,]
  summary(tal.ni)

tal.sem <- lm(REL_TAL~log(SEMNi.surf),data=chem.inv)
  summary(tal.sem)

tal.semavs <- lm(REL_TAL~log(SEMAVSfOC.avg),data=chem.inv)
  summary(tal.semavs)

tal.semhfo <- lm(REL_TAL~log(NiSEMHFOfOC),data=chem.inv)
  summary(tal.semhfo)
  #abline(tal.semhfo)


par(pin=c(3,3))

##TOTAL Ni SURFACE || Cannot fit sigmoid curve

plot(REL_TAL~TotalNi.surf.mol,log="x",data=chem.inv[1:20,], xlim=c(0.5,10), ylim=c(0,45), cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("Ni"[TOT]*" (µmol g"^"-1"," "[DW]*")")),
     ylab="Talitridae relative abundance (%)") 
      #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_TAL~TotalNi.surf.mol,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topleft", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_TAL~a/(1+10^(2*b*(log10(TotalNi.surf.mol)-c))),start=list(a=50,b=1,c=1),data=chem.inv)
CRMNi=seq(-2,3,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(TotalNi.surf.mol=CRMNi))
lines(CRMNi,CRMy,lwd=3)

##SEM Ni || Cannot fit sigmoid curve

plot(REL_TAL~SEMNi.surf.mol,log="x",data=chem.inv[1:20,],cex=1.5, xlim=c(0.2,10), ylim=c(0,45), col=2,
     pch=16,las=1,xlab=expression(paste("Ni"[SEM]*" (µmol g"^"-1"," "[DW]*")")),ylab="Talitridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_TAL~SEMNi.surf.mol,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topleft", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)


##SEM-AVS/ƒOC (C.avg) || Model fit looks pretty awful

plot(REL_TAL~SEMAVSfOC.avg,log="x",data=chem.inv[1:20,], xlim=c(5,120), ylim=c(0,45),cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("(Ni"[SEM]*"-AVS)/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),ylab="Talitridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_TAL~SEMAVSfOC.avg,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topleft", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_TAL~a/(1+10^(4*b*(log10(SEMAVSfOC.avg)-c))),start=list(a=50,b=1,c=2),data=chem.inv)
CRMNi=seq(0.01,3,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(SEMAVSfOC.avg=CRMNi))
lines(CRMNi,CRMy,lwd=3)


##SEM Ni-HFO/ƒOC

plot(REL_TAL~NiSEMHFOfOC,log="x",data=chem.inv[1:20,],xlim=c(2,100), ylim=c(0,45), cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("(Ni"[SEM]*"-Ni"[HFO]*")/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),
     ylab="Talitridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_TAL~NiSEMHFOfOC,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topleft", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_TAL~a/(1+10^(3*b*(log10(NiSEMHFOfOC)-c))),start=list(a=50,b=1,c=2),data=chem.inv)
CRMNi=seq(0.01,4,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(NiSEMHFOfOC=CRMNi))
lines(CRMNi,CRMy,lwd=3)
```

_____

####7. Sigmoid CRM EPH
```{r 7. Sigmoid CRM EPH, echo=F, fig.align='center',tidy=T}
par(pin=c(3,3))

chem.inv$TotalNi.surf.mol = chem.inv$TotalNi.surf/58.69
chem.inv$SEMNi.surf.mol = chem.inv$SEMNi.surf/58.69

##TOTAL Ni SURFACE (Molar)

plot(REL_EPH~TotalNi.surf.mol,log="x",data=chem.inv[1:20,], xlim=c(0.5,10), ylim=c(0,45), cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("Ni"[TOT]*" (µmol g"^"-1"," "[DW]*")")),
     ylab="Ephemeridae relative abundance (%)") 
      #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_EPH~TotalNi.surf.mol,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topright", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_EPH~a/(1+10^(2*b*(log10(TotalNi.surf.mol)-c))),start=list(a=50,b=0.5,c=1),data=chem.inv)
CRMNi=seq(-2,3,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(TotalNi.surf.mol=CRMNi))
lines(CRMNi,CRMy,lwd=3)

  #Standard error for all predicted values
#se.pred = sqrt(apply(attr(predict(CRM, data.frame(TotalNi.surf.mol=CRMNi)),"gradient"),1,function(x) sum(vcov(CRM)*outer(x,x))))
  #Lower bound of 95% confidence intervals (2.5%)
#low.CI=CRMy+se.pred*qnorm(0.025)
  #Upper bound of 95% confidence intervals (97.5%)
#up.CI=CRMy+se.pred*qnorm(0.975)

#lines(CRMNi,CRMy,lwd=3,col="red")
#lines(CRMNi,low.CI,lwd=2,lty=2)
#lines(CRMNi,up.CI,lwd=2,lty=2)

CRM=nls(REL_EPH~SSlogis(log10(TotalNi.surf.mol),Asym,xmid,scal),data=chem.inv)
summary(CRM) #EC20 = 151.4 mg/kg Total Ni // EC20 = 2.57 µmol/g Total Ni

ANOVA_F(CRM,chem.inv$REL_EPH)

#Set the reference growth rate
ref = CRM$m$getPar()[1]
  #For EC20
CRM.EC20=which.min(abs(CRMy-ref*0.8))
(EC20 = CRMNi[CRM.EC20])
ECXX(CRM,20)


##SEM Ni (Molar)

plot(REL_EPH~SEMNi.surf.mol,log="x",data=chem.inv[1:20,],cex=1.5, xlim=c(0.2,10), ylim=c(0,45), col=2,
     pch=16,las=1,xlab=expression(paste("Ni"[SEM]*" (µmol g"^"-1"," "[DW]*")")),ylab="Ephemeridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_EPH~SEMNi.surf.mol,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topright", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_EPH~a/(1+10^(2*b*(log10(SEMNi.surf.mol)-c))),start=list(a=50,b=0.5,c=1),data=chem.inv)
CRMNi=seq(-2,3,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(SEMNi.surf.mol=CRMNi))
lines(CRMNi,CRMy,lwd=3)

CRM=nls(REL_EPH~SSlogis(log10(SEMNi.surf.mol),Asym,xmid,scal),data=chem.inv)
summary(CRM) #EC20 = 87.1 mg/kg SEM Ni // EC20 = 1.51 µmol/g SEM Ni

ANOVA_F(CRM,chem.inv$REL_EPH)

#Set the reference growth rate
ref = CRM$m$getPar()[1]
  #For EC20
CRM.EC20=which.min(abs(CRMy-ref*0.8))
(EC20 = CRMNi[CRM.EC20])
ECXX(CRM,20)


##SEM-AVS/ƒOC (C.avg) // (SEM-AVS corrected for average C from reference and exposure surface sediment)

plot(REL_EPH~SEMAVSfOC.avg,log="x",data=chem.inv[1:20,], xlim=c(5,120), ylim=c(0,45),cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("(Ni"[SEM]*"-AVS)/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),ylab="Ephemeridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_EPH~SEMAVSfOC.avg,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topright", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_EPH~a/(1+10^(2*b*(log10(SEMAVSfOC.avg)-c))),start=list(a=50,b=1,c=2),data=chem.inv)
CRMNi=seq(0.01,3,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(SEMAVSfOC.avg=CRMNi))
lines(CRMNi,CRMy,lwd=3)

CRM=nls(REL_EPH~SSlogis(log10(SEMAVSfOC.avg),Asym,xmid,scal),data=chem.inv)
summary(CRM) #EC20 = 26.9 umol/g SEM-AVS/ƒOC (avg)

ANOVA_F(CRM,chem.inv$REL_EPH)

#Set the reference growth rate
ref = CRM$m$getPar()[1]
  #For EC20
CRM.EC20=which.min(abs(CRMy-ref*0.8))
(EC20 = CRMNi[CRM.EC20])
ECXX(CRM,20)


##SEM Ni-HFO/ƒOC (Molar)

plot(REL_EPH~NiSEMHFOfOC,log="x",data=chem.inv[1:20,],xlim=c(2,100), ylim=c(0,45), cex=1.5, col=2,
     pch=16,las=1,xlab=expression(paste("(Ni"[SEM]*"-Ni"[HFO]*")/",ƒ,""[OC]*" (µmol g"^"-1"," "[OC]*")")),
     ylab="Ephemeridae relative abundance (%)") #Ref values are all the average of the 4 ref sites from the respective mine
points(REL_EPH~NiSEMHFOfOC,data=chem.inv[21:40,],cex=1.5,pch=16,las=1, xlab="n",ylab="n")

legend("topright", c("Birch", "Thompson"), text.col=c(2,1), bty="n", text.font=2)

CRM=nls(REL_EPH~a/(1+10^(3*b*(log10(NiSEMHFOfOC)-c))),start=list(a=50,b=1,c=1.5),data=chem.inv)
CRMNi=seq(0.01,4,by=0.01)
CRMNi=10^CRMNi
CRMy=predict(CRM,list(NiSEMHFOfOC=CRMNi))
lines(CRMNi,CRMy,lwd=3)

CRM=nls(REL_EPH~SSlogis(log10(NiSEMHFOfOC),Asym,xmid,scal),data=chem.inv)
summary(CRM) #EC20 = 891.3 mg/kg SEM-HFO/ƒOC (avg) // EC20 = 15.49 µmol/g SEM-HFO/ƒOC (avg)

ANOVA_F(CRM,chem.inv$REL_EPH)

  #Set the reference growth rate
ref = CRM$m$getPar()[1]
  #For EC20
CRM.EC20=which.min(abs(CRMy-ref*0.8))
(EC20 = CRMNi[CRM.EC20])
ECXX(CRM,20)
```

|*Log-Logistic* Model|EC20 (µmol/g)|
|:-:|:-:|
|Total Ni|2.6|
|SEM Ni|1.5|
|SEM-AVS/ƒOC (w/ 'outliers')|26.9|
|SEM-HFO/ƒOC|15.5|